---
title: "Assignment 1 - The tidyverse"
author: "Adrian Werner"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    number_sections: no
    theme: cosmo
    toc: no
  pdf_document:
    toc: no
---
  
<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

<!-- Do not forget to input your Github username in the YAML configuration up there --> 

***

```{r, include = T}
library("tidyverse")
library(legislatoR)
library(lubridate)
```

<br>

### Getting started with the Comparative Legislators Database

The Comparative Legislators Database (CLD) includes political, sociodemographic, career, online presence, public attention, and visual information for over 45,000 contemporary and historical politicians from ten countries (see the [corresponding article](https://www.cambridge.org/core/journals/british-journal-of-political-science/article/comparative-legislators-database/D28BB58A8B2C08C8593DB741F42C18B2) for more information). It can be accessed via `legislatoR` - an R package that is available on both [CRAN](https://cran.r-project.org/web/packages/legislatoR/index.html) and [GitHub](https://github.com/saschagobel/legislatoR).

Before you start with the following tasks, skim the tutorial to make yourself familiar with the database. You find it [here](https://cran.r-project.org/web/packages/legislaaut_house_core4toR/vignettes/legislatoR.html).

For the following tasks, you will work with ONE of the legislatures. The legislature you'll work with depends on your first name:

| Your first name starts with... | Legislature | Code |
|---------|-------|-------|
| A-C | Austrian Nationalrat | `aut` |
| D-F | Canadian House of Commons | `can` |
| G-I | Czech Poslanecka Snemovna | `cze` |
| J-L | Spanish Congreso de los Diputados | `esp` |
| M-O | French Assemblée | `fra` |
| P-R | German Bundestag | `deu` |
| S-U | Irish Dail | `irl` |
| V-X | UK House of Commons | `gbr` |
| Y-Z | US House of Representatives  | `usa_house` |

The following tasks will be based on data from the `core` and `political` tables of the database.

<br>

***

### Task 1 - Descriptives [8 points in total]

a) What's the overall share of male legislators in the entire dataset? [1 point]

```{r}
aut_house_core <- get_core(legislature = "aut")

table(aut_house_core$sex)/length(aut_house_core$sex)
```

b) How many (both in absolute and relative terms) legislators died in the same place they were born in? [1 point]

```{r}

aut_house_core %>%
  filter(aut_house_core$birthplace == aut_house_core$deathplace) %>%
  nrow()
 
aut_house_core %>%
  filter(aut_house_core$birthplace == aut_house_core$deathplace) %>%
  nrow()/nrow(aut_house_core)   

```

c) Create a new variable `age_at_death` that reports the age at which legislators died. Then, plot the distribution of that variable for the entire dataset. [2 points]

```{r}
aut_house_core2 <- aut_house_core %>%
  select(name, death, birth) %>%
  mutate(
    age_at_death = time_length(difftime(aut_house_core$death, aut_house_core$birth), "years")
  )

hist(aut_house_core2$age_at_death, main = "Distribution Age of Death", xlab = "Age")
    
```

d) What is the most frequent birthday in your sample (e.g., “January 15")? [2 points]

```{r}
birth_fac <- as.factor(aut_house_core$birth)
names(which.max(table(birth_fac)))
```

e) Generate a table that provides the 5 legislators with the longest names (in terms of number of characters, ignoring whitespace). [2 points]

```{r}
aut_house_core3 <- aut_house_core %>%
  select(name) %>%
  mutate(
    character_length = nchar(name))  %>% 
  arrange(desc(character_length)
  )
  
 aut_house_core4 <- head(aut_house_core3, 5)
 
 aut_house_core4
```

<br>

***

### Task 2 - Exploring trends in the data [8 points in total]

a) Using data from all sessions, provide a plot that shows the share of male legislators by session! [3 points]

```{r}

aut_house_pol <- get_political(legislature = "aut")

merged_df <- merge.data.frame(aut_house_core, aut_house_pol)

merged_df %>% 
  mutate(gender = ifelse(merged_df$sex == "male", 1, 0)) %>%
  group_by(session) %>%
    summarise(perc_male = ((sum(gender)/length(gender))))
```

b) Explore another relation of variables in the dataset. Your analysis should (i) use data from both the `core` and the `political` table and (ii) feature data from several sessions. Also, offer a brief description and interpretation of your findings! [5 points]

```{r}

merged_df %>%
  mutate(death_no = ifelse(is.na(merged_df$death), 0, 1)) %>%
  group_by(party) %>%
    summarise(death_no = sum(death_no)/length(death_no)) %>%
  arrange(desc(death_no))

```

<div class = "answer">
The table shows, in absolute terms, the death rate among party members in a descending order. Interestingly, there are a number of parties with a death rate of 100%. This is because they are members of parties that do not exist any longer. These parties are therefore not only formally but also in terms of their representatives vanished from the political landscape. Interestingly, even the biggest parties SPÖ (founded 1889) and ÖVP (founded 1945) are >50% "dead", indicating that Austria is already a old democracy.    
</div>

<br>

***



